### 2017-05-17 sheets/basis-3/sheet-basis-3.tex
Template for exercise sheets
Add exercises for the third sheet

% Use Basis x or Talk x, where x is the number of the session
\uebunghead{Basis 2}{Maps between Categories -- Functors}

\begin{hint}
  Read Chapter 1.2.
\end{hint}

\section{Diagrams}

\begin{hint}
  This is how you can draw diagrams:
  \[
    \begin{tikzcd}
                                    & A \arrow{dr}{g}    & \\
      B \arrow{ur}{f} \arrow{rr}{h} &                    & C
    \end{tikzcd}
  \]
\end{hint}

>>>>

\author{Joachim Bard, Maximilian Wuttke, Nikita Ziuzin}

% Use Basis x or Talk x, where x is the number of the session
\uebunghead{Basis 3}{Maps between Functors -- Natural Transformations}

\begin{hint}
  Read Chapter 1.3.
\end{hint}

\begin{exercise}
	Let $\cat{A}$ and $\cat{B}$ be categories.  Prove that
	$\ftrcat{\cat{A}^\op}{\cat{B}^\op} \iso \ftrcat{\cat{A}}{\cat{B}}^\op$.
\end{exercise}

\begin{exercise}

  \begin{definition}{Canonical according to Leinster}
     "Canonical is an informal word, meaning something like 'God-given' or
    'defined without making arbitrary choices'.  For example, for any two sets
    $A$ and $B$, there is a canonical bijection $A \times B \to B \times A$
    defined by $(a, b) \mapsto (b, a)$, and there is a canonical function $A
    \times B \to A$ defined by $(a, b) \mapsto a$.  But the function $B \to A$
    defined by 'choose an element $a_0 \in A$ and send everything to $a_0$' is
    not canonical, because the choice of $a_0$ is arbitrary."
  \end{definition}
	Let $A$ and $B$ be sets, and denote by $B^A$ the set of functions from $A$
	to $B$.  Write down:
	\begin{itemize}
		\item[(a)] a canonical function $A \to B^{(B^A)}$,
		\item[(b)] a canonical function $A \times B^{A} \to B$.
	\end{itemize}
\end{exercise}

\begin{exercise}
  For a fixed set $A$, the map taking $B$ to $B^A \times A$ can be extended to
  a functor $F_A:\mathbf{Set} \to \mathbf{Set}$ as follows:
  \begin{align*}
    F_A(B) & =  B^A \times A\\
    F_A(f: B \to C) & =  (F \circ \_) \times id_A
  \end{align*}
  The fact that $eval: F_A \to I_{\mathbf{Set}}$ is a natural transformation follows from the commutativity of the diagram
\[
  \begin{tikzcd}
    F_A(C) = C^A \times A \arrow{rrr}{eval_{AC}} \arrow{dd}{F_A(g)=(g\circ \_) \times id_A} & & & C = I_{\mathbf{Set}}(C) \arrow{dd}{g = I_{\mathbf{Set}(g)}} \\
                                                                                           & & & \\
    F_A(B) = B^A \times A \arrow{rrr}{eval_{AB}} & & & B = I_{\mathbf{Set}}(B)
  \end{tikzcd}
\]
for each $g: C \to B$.

Check that this diagram commutes.
\end{exercise}

\begin{exercise}
	Let $\xymatrix@1{\cat{A}\rtwocell^F_G{\alpha} &\cat{B}}$ be a natural transformation.
	\begin{itemize}
		\item[(a)] Suppose $\alpha$ is a natural isomorphism.
		Show that $\alpha_A\from F(A) \to G(A)$ is an isomorphism for all $A \in \cat{A}$.
		\item[(b)] Now assume that $\alpha_A\from F(A) \to G(A)$ is an isomorphism for all
		$A \in \cat{A}$. Prove that $\alpha$ is a natural isomorphism.
	\end{itemize}
\end{exercise}

\begin{exercise}
	Let the following natural transformations be given:
	\[
\xymatrix{
\cat{A}
\ruppertwocell<8>^F{\alpha}
\ar[r]|G
\rlowertwocell<-8>_H{\beta} &
\cat{A}'
\ruppertwocell<8>^{F'}{\hspace{.2em}\alpha'}
\ar[r]|{G'}
\rlowertwocell<-8>_{H'}{\hspace{.2em}\beta'} &
\cat{A}''
}
\]
Now prove that the interchange law holds, that is
\[
(\beta' \of \alpha') * (\beta \of \alpha)
=
(\beta' * \beta) \of (\alpha' * \alpha)
\]
\end{exercise}

\begin{exercise}
A \emph{permutation} of a set $X$ is a bijection $X \to X$.
Write $\Sym(X)$ for the set of permutations of $X$.
A \emph{total order} on a set $X$ is an order $\leq$ such that for all $x, y \in X$,
either $x \leq y$ or $y \leq x$.
Write $\Ord(X)$ for the set of total orders on $X$.

Let $\cat{B}$ denote the category of finite sets and bijections.

	\begin{itemize}
	\item[(a)]
Give a definition of $\Sym$ on maps in $\cat{B}$ in such a way that $\Sym$
becomes a functor $\cat{B} \to \Set$.  Do the same for $\Ord$.  Both your
definitions should be canonical (no arbitrary choices).

	\item[(b)]
Show that there is no natural transformation $\Sym \to \Ord$.
(Hint: consider identity permutations.)

	\item[(c)]
For an $n$-element set $X$, how many elements do the sets $\Sym(X)$
and $\Ord(X)$ have?
	\end{itemize}

Conclude that $\Sym(X) \iso \Ord(X)$ for all $X \in \cat{B}$, but not
\emph{naturally} in $X \in \cat{B}$.  (The moral is that for each finite
set $X$, there are exactly as many permutations of $X$ as there are total
orders on $X$, but there is no natural way of matching them up.)
\end{exercise}

\begin{exercise}
Let $F\from \cat{A} \to \cat{B}$ be a functor.

	\begin{itemize}
	\item[(a)]
Suppose that $F$ is an equivalence.  Prove that $F$ is full, faithful and
essentially surjective on objects.  (Hint: prove faithfulness before
fullness.)

	\item[(b)]
Now suppose instead that $F$ is full, faithful and essentially surjective
on objects.  For each $B \in \cat{B}$, choose an object $G(B)$ of $\cat{A}$
and an isomorphism $\varepsilon\from F(G(B)) \to B$.  Prove that $G$ extends
to a functor in such a way that $(\varepsilon)_{B \in \cat{B}}$ is a natural
isomorphism $F \circ G \to 1_{\cat{B}}$.  Then construct a natural isomorphism
$1_{\cat{A}} \to G \circ F$, thus proving that $F$ is an equivalence.
	\end{itemize}
\end{exercise}

\begin{exercise}
This exercise makes precise the idea that linear algebra can equivalently
be done with matrices or with linear maps.

Fix a field $K$.  Let $\Mt$ be the category whose objects are the natural
numbers and with
\[ \Mt(m, n) = \{ n \times m \text{ matrices over } K \}. \]
Prove that $\Mt$ is equivalent to $\FDVect_K$, the category of
finite-dimensional vector spaces over $K$.
Does your equivalence involve a \emph{canonical} functor from $\Mt$ to $\FDVect_K$, or from $\FDVect_K$ to $\Mt$?
\end{exercise}
